---
title: "Practical Machine learning assessment"
output: 
  html_document:
    keep_md: true
---
#INTRODUCTION
This project is about predicting the manner in which the subjects performed weight lifting exercise. The data is collected from devices such as Jawbone Up, Nike FuelBand, and Fitbit.

##Data Preprocessing
Set the working driectory and load the required libraries and set seed for results to be Reproducible.
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
setwd("C:\\project_8")
library(caret)
library(randomForest)
set.seed(50)
```
Load the given csv file to the variables
```{r}
pml_train <- read.csv("pml-training.csv",header=T,sep=",",na.strings=c("NA","NaN", " ", "", "#DIV/0!"))
pml_test <- read.csv("pml-testing.csv",header=T,sep=",",na.strings=c("NA","NaN", " ", "", "#DIV/0!"))
```
Removing the first column that represents the ID
```{r}
pml_train <- pml_train[,-1]
pml_test <- pml_test[,-1]
```
##Cleansing the data for the model ( We are using Random Forest method here)
Remove columns that do not satisfy is applied before applying to the model
```{r}
remove_train <- c((colSums(!is.na(pml_train[,-ncol(pml_train)])) >= 0.6*nrow(pml_train)))
pml_train   <-  pml_train[,remove_train]
```
Create data partitions for training and validation
```{r}
trainPart <- createDataPartition(pml_train$classe, p=0.6, list=FALSE)
pml_training <- pml_train[trainPart,]
pml_validation <- pml_train[-trainPart,]
```
Remove the near zero variance columns
```{r}
Train_NonZeroVar <- nearZeroVar(pml_training)
Validate_NonZeroVar <- nearZeroVar(pml_validation)
pml_train_without_nzv <- pml_training[, -Train_NonZeroVar]
pml_validate_without_nzv <- pml_validation[,-Validate_NonZeroVar]
```
##Building the model
```{r}
PredictModel <- randomForest(classe~.,data=pml_train_without_nzv,importance=TRUE)
PredictModel
```
Now evaluate the results through the confusion matrix
```{r}
confusionMatrix(predict(PredictModel,newdata=pml_validate_without_nzv[,-ncol(pml_validate_without_nzv)]),
                pml_validate_without_nzv$classe)
```
Check the accuracy of the model
```{r}
accuracy<-c(as.numeric(predict(PredictModel,newdata=pml_validate_without_nzv
                            [,-ncol(pml_validate_without_nzv)])==pml_validate_without_nzv$classe))
accuracy<-sum(accuracy)*100/nrow(pml_validate_without_nzv)
accuracy
```
We can see our model is having an accuracy of 99.89 % and the out-of-sample error is only 0.11%

##Testing the model
Clean the test data as same as like we did for training set
```{r}
remove_test <- c((colSums(!is.na(pml_test[,-ncol(pml_test)])) >= 0.6*nrow(pml_test)))
pml_test   <-  pml_test[,remove_test]
pml_test <- pml_test[,-ncol(pml_test)]
```
Making the testing dataset to same class and structure of training dataset and apply dummy row id for dummy in the test data
```{r}
pml_test <- rbind(pml_train[100, -59] , pml_test)
row.names(pml_test) <- c(100,1:20)
```
Testing the model with the test data
```{r}
Prediction<- predict(PredictModel,newdata=pml_test[-1,])
Prediction
```